Democracy, Public Support, and Measurement Uncertainty

### 2021-08-08 paper/dcpo_demsupport.Rmd
update moc and corresponding bib

Although measurement uncertainty has yet attracted attention in the field of comparative public opinion [see, in addition to the works examined here, @OGrady2019], latent variables are estimated with a quantifiable amount of measurement uncertainty, and ignoring that uncertainty in analyses can bias coefficient estimates and standard errors.
Especially when measurement error is correlated with the outcome variable, the resulting bias could either attenuate or exaggerate the magnitude of the estimated effects [see, e.g., @Blackwell2017, p.360; @caughey2018, p.254].
In light of this, recent studies measuring other latent variables have recommended incorporating their measurement uncertainty in analyses [see @solis_measuring_2020, p. 18; @Gandhi2020, p. 1553], and research examining the consequences of public opinion in the United States has done so [see, e.g., @Kastellec2015, pp. 791-792; @caughey2018, p. 254]. 
Following this line, after replicating the original analyses with point estimates, we incorporate uncertainty in analysis models using our replication data.
We take the distribution of the latent (explanatory) variable, the outcome variable, and control variables to account for the uncertainty.
We take 900 draws from the posterior distribution of democratic support, outcome, and control variables with errors.
We then duplicate each article's analysis dataset of variables "without" measurement error 900 times, and assign them to each a different random draw from the distributions of variables with measurement error. 
We conduct inferences from the distributive data via a technique known as "method of composition" (MOC) [@Treier2008, p. 215].
See more details of the method and the operational process in Online Supplementary Material \@ref(moc).

@@

# Replication with Alternative Methods Accounting for Uncertainty {#moc}

To ensure the robustness of our analysis, we apply another method to account for uncertainty.
The method we employ is "method of composition" (MOC).
It is a common method for Monte Carlo integration, but can also be applied for addressing uncertainty [@Treier2008].
To compare these two methods, let's denote `y` as a vector of the outcome variable [either institutional democracy in @Claassen2020a or the changes in mass support of democracy in @Claassen2020b]. 
Let $x$ (variables without subscript are used to denote vectors of observations, viz., $x$ = ($x_1$,......, $x_i$)) be a vector of the explanatory variable (e.g., mass support or institutional democracy) and $\theta$ be the parameter of interest.

Rubin's Rules (RR, the method used in the main article) take variables with measurement error as missing value, and combine the point estimates and standard errors into single values yet accounting for both within and between imputation variances [@Rubin1987].

Rubein's Rules is defined in the following formulas.

\begin{equation}
\bar{\theta}_{m} = \frac{1}{m}\sum_{i=1}^{m} \theta_{i}.
\end{equation}

\begin{equation}
{V}_{B} = \frac{1}{m-1}\sum_{i=1}^{m} (\theta_{i}-\bar{\theta}_{m})^2.
\end{equation}

\begin{equation}
{V}_{W} = \frac{1}{m}\sum_{i=1}^{m} SE_{i}^2.
\end{equation}

\begin{equation}
{V}_{T} = {V}_{w} +(1+ \frac{1}{m})\frac{V_B}{m}.
\end{equation}

$\bar{\theta}_{M}$ is the pooled parameter estimate, $m$ is the number of imputed datasets, $V_{B}$ is between imputation variance, $V_{W}$ within imputation variance, $V_{T}$ is total variance.

The general formula for confidence interval (CI) is defined as:

\begin{equation}
\bar{\theta} \pm t_{df,1-\alpha/2} *\sqrt{V_{T}}.
\end{equation}

$df$ is calculated by using $df_{Adjusted}$ [@barnard1999] and the Fraction of Missing information, lambda, is derived from between and total variance [@raghunathan2015, @van2018].

\begin{equation}
{df_{Adjusted}} = \frac{df_{Old}*df_{Observed}}{df_{Old}+df_{Observed}}.
\end{equation}

\begin{equation}
{df_{Old}} = \frac{(n-k)+1}{(n-k)+3} * (n-k)(1-lambda).
\end{equation}

\begin{equation}
{df_{Observed}} = \frac{(n-k)+1}{(n-k)+3} * (n-k)(1-lambda).
\end{equation}

\begin{equation}
{Lambda} = \frac{V_{B}+\frac{V_{B}}{m}}{V_{T}}
\end{equation}

One main idea behind RR is to account for both within imputation variance and sample variance which is due to the missing data (measurement error in our case).

In analysis models, we have latent variables in both sides of equations, public suppport for democracy, democracy indices and corruption index (The latter one is included in APSR analyses).
Since measurement uncertainty associated with these latent variables can propogate into the inferences over coefficient parameters in models, we incorporate uncertainty by employing "Method of Composition" [@Tanner1996,p52], which has been applied in analyses with latent variables [@Trei2008;@Caughey2018;@K2015].
As @Caughey2018 explained, the main idea of MOC is to estimate the marginal distribution of coefficient parameter vector $\beta$, intergrating over the uncertainty in latent variables $\theta$ (P A-15).
More explicitly, we wish to integrate the joint density of $\beta$ and $\theta$ over the distribution of $\theta$.

@@

We incorporate uncertainty for five variables, public support for democracy, liberal democracy, electoral democracy, liberal component index, and corruption perceptions index.

MOC addresses the same issue from a slightly different angle.
The method assumes that the explanatory and outcome variables are samples from their estimate functions, and their paired sample follows a joint density [@Tanner1993, p.31].
From the joint density, one can decomposite the likelihood of the explanatory variable and its parameter to a product of their own probabilities.
When x is measured with errors $w$, the probability becomes conditional:

\begin{equation}
\mathcal{L}(\theta, x|y, w) = p(\theta|x,y)p(x|w).
\end{equation}

Based on this decomposition, one can estimate the marginal density for &\theta&:

\begin{equation}
\mathcal{L}(\theta|y, x) \propto \int_{\chi} p(\theta|y,x)p(x|w)dx.
\end{equation}

Again, the gained parameter density can be combined into single values.
A common way to do so is using the mean of $theta$ as the point estimate, and the variance from the density as the standard error.

>>>>

Although measurement uncertainty has yet attracted attention in the field of comparative public opinion [see, in addition to the works examined here, @OGrady2019], latent variables are estimated with a quantifiable amount of measurement uncertainty, and ignoring that uncertainty in analyses can bias coefficient estimates and standard errors.
Especially when measurement error is correlated with the outcome variable, the resulting bias could either attenuate or exaggerate the magnitude of the estimated effects [see, e.g., @Blackwell2017, p.360; @caughey2018, p.254].
In light of this, recent studies measuring other latent variables have recommended incorporating their measurement uncertainty in analyses [see @solis_measuring_2020, p. 18; @Gandhi2020, p. 1553], and research examining the consequences of public opinion in the United States has done so [see, e.g., @Kastellec2015, pp. 791-792; @caughey2018, p. 254]. 
Following this line, after replicating the original analyses with point estimates, we incorporate uncertainty in analysis models using our replication data.
We conduct inferences from the distributive data via a technique known as "method of composition" (MOC) [@Tanner1996, p.52;@Treier2008, p. 215; @caughey2018,p.A15-16].
See more details of the method and the operational process in Online Supplementary Material \@ref(moc).

@@

# Replication with MOC Accounting for Uncertainty {#moc}

In analysis models, we have latent variables in both sides of equations, public suppport for democracy, democracy indices and corruption index (The latter one is included in APSR analyses).
Since measurement uncertainty associated with these latent variables can propogate into the inferences over coefficient parameters in models, we incorporate uncertainty by employing "Method of Composition" [@Tanner1996,p52], which has been applied in analyses with latent variables [@Treier2008;@Caughey2018;@Kastellec2015].

As @Caughey2018 explained, the main idea of MOC is to estimate the marginal distribution of coefficient parameter vector $\beta$, intergrating over the uncertainty in latent variables $\theta$ (P A-15).
More explicitly, we wish to integrate the joint density of $\beta$ and $\theta$ over the distribution of $\theta$.

@@

We incorporate uncertainty for five variables, public support for democracy, liberal democracy, electoral democracy, liberal component index, and corruption perceptions index.
For each of these five latent variables, we take 900 draws from its posterior distribution. We duplicate the dataset of variables “without” measurement error 900 times, and assign them to each a different random draw from the distributions of variables with measurement error, which yields 900 dataset.
In next step, we run each of analysis models with these 900 datesets independently and save the resulting estimates of coefficients and the matrix of variance-covariance for each run.
We then draw one sample from the mulitivariate normal distribution with the mean vector of coefficient estimates and variance-covariance matrix produced from each run.
This procedure finally yields 900 samples of estimated coefficients drawn from the joint density of $\beta$ and $\theta$.
We calcluate point estimates and standard error based on these 900 samples.
